---
title: "Bicycle Sharing Activity"
author: "Kashish Gujral"
due date: "Oct 18, 2020"
output: html_notebook
---

```{r}
#load packages
library(mosaic)
library(tidyverse)
library(lubridate)
library(devtools)
library(leaflet)
```

#### Load Data

```{r}
Stations <- readr::read_csv("http://tiny.cc/dcf/DC-Stations.csv")
data_site <- "http://tiny.cc/dcf/2014-Q4-Trips-History-Data-Small.rds"
    Trips <- readRDS(gzcon(url(data_site)))
```

```{r}
Trips %>%
  ggplot(aes(x=sdate)) +
  geom_density(fill="gray", color=NA)
```
# 18.5 How long?

#### The distribution of the duration of rental events, broken down by the client type.
```{r}
Trips <- Trips %>%
  mutate(duration = as.numeric(edate - sdate)/60)
```

```{r}
Trips %>%
  ggplot(aes(x = client, y =duration)) +
  geom_boxplot() +
  labs(x = "Client Type" , y = "Rental Duration In Minutes") +
  ylim(0,60)
           
```

# 18.6 When are bikes used?

```{r}
Trips <-Trips %>%
  mutate(dayOfYear = lubridate::yday(sdate),dayOfWeek = lubridate::wday(sdate),dayOfWeekLabel = lubridate::wday(sdate, label = TRUE),hourOfDay = lubridate::hour(sdate),minuteOfHour = lubridate::minute(sdate))
```


#### Hours

#### The graph below shows distribution of bike trips by hour of the day.Few bikes are checked out before 5am, and that there are busy times around the rush hour: 8am and 5pm.
```{r}
Trips %>%
  mutate(H = lubridate::hour(sdate)) %>%
  ggplot(aes(x = H)) +
  geom_density(fill = "grey" , adjust = 2) +
  labs(x = "Hours")
  
```

#### Minutes

#### The graph below shows distribution of bike trips by Minutes in the day.
```{r}
Trips %>%
  mutate(M = lubridate::minute(sdate)) %>%
  ggplot(aes(x = M)) +
  geom_density(fill = "grey" , adjust = 2) +
  labs(x = "Minutes")
```
#### Day of the Year

#### The graph below shows distribution of bike trips by Day of the Year.
```{r}
Trips %>%
  mutate(yday = lubridate::yday(sdate)) %>%
  ggplot(aes(x = yday)) +
  geom_density(fill = "grey" , adjust = 2) +
  labs(x = "Day of year")
```
#### Day of the Week

#### The graph below shows distribution of bike trips by Day of the week. few bikes are checked out during the weekends, and that there are busy times around the weekdays.
```{r}
Trips %>%
  mutate(wday = lubridate::wday(sdate, label = T)) %>%
  ggplot(aes(x = wday)) +
  geom_bar(fill = "grey" , adjust = 2) +
  labs(x = "Day of the Week")

```
#### Group rental by hour, day of the week, and client type

#### Frequency
```{r}
Trips%>%
  group_by(client, dayOfWeek, hourOfDay) %>%
  summarise(count = n())%>%
  ggplot(aes(x = hourOfDay , y = count)) +
  geom_line(aes(group = dayOfWeek, color = as.character(dayOfWeek))) +
  facet_wrap(~client)
```
#### Density

```{r}
Trips%>%
  ggplot(aes(x = hourOfDay)) +
  geom_density(aes(group = dayOfWeek, color = as.character(dayOfWeek))) +
  facet_wrap(~client)
```

# 18.7 How Far?

```{r}
source("http://tiny.cc/dcf/haversine.R")
```

```{r}
Simple <-
  Stations %>%
  select(name, lat, long) %>%
  rename(sstation=name)
Simple2 <-
  Simple %>%
  rename(estation=sstation, lat2=lat, long2=long)
```

```{r}
StationPairs <- merge(Simple, Simple2, by=NULL)

```

```{r}
PairDistances <-
  StationPairs %>%
  mutate(distance = haversine(lat, long, lat2, long2)) %>%
  select(sstation, estation, distance)
```

```{r}
RideDistances <-
  Trips %>%
  inner_join(PairDistances)
```

```{r}
RideDistances %>%
  ggplot(aes(x= distance))+
  geom_density(fill = "red")+
  geom_density(data = PairDistances, aes(x = distance) , fill = "blue" , alpha = 0.7)
```

# 18.8 Mapping Stations


```{r}
stationMap <-
  leaflet(Stations) %>%   # like ggplot()
  addTiles() %>%          # add the map
  addCircleMarkers(radius=2, color="red") %>%
  setView(-77.04, 38.9, zoom=12)
```

```{r}
stationMap
```
# 18.9 Long-Distance stations(Extra Credit)

```{r}
StationPairmap <-
  RideDistances%>%
  rename(name = sstation) %>%
  left_join(Stations) %>%
  transmute(name, lat, long , disProxy = distance * 1000/3) %>%
  group_by(name, lat, long) %>%
  summarise(medDis = median(disProxy))
```

```{r}
head(StationPairmap)
```

```{r}
distancemap <-
  leaflet(StationPairmap) %>%
  addTiles() %>%    
  addCircleMarkers(radius=2, color="red") %>%
  addCircles(radius = ~ mid, color = "blue", opacity = 0.0001)
  setView(-77.04, 38.9, zoom=12)
```

```{r}
distancemap
```


```{r}
freqmap <-
  leaflet(StationPairmap) %>%   # like ggplot()
  addTiles() %>%          # add the map
  addCircleMarkers(radius=2, color="red") %>%
  addCircles(radius= ~medDis, color="blue", opacity = 0.0001)
  setView(-77.04, 38.9, zoom=12)
```

```{r}
freqmap
```

